---
title: "HW 4"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE,  
                      warning = FALSE, message = FALSE, 
                      fig.align = "center",
                      R.options = list(max.print=100))

# Edit the file starting below
```

### Enter your name and EID here: Callihan Bertley cjb5564

**You will submit this homework assignment as a HTML file on Canvas.**

*For all questions, include the R commands/functions that you used to find your answer (show R chunk). Answers without supporting code will not receive credit. Write full sentences to describe your findings.*

------------------------------------------------------------------------

### Question 1: (2 pts)

All subsequent code will be done using `dplyr`, so we need to load this package. We also want to look at the `penguins` dataset which is inside the `palmerpenguins` package:

```{r message=FALSE}
# Call dplyr and ggplot2 packages within tidyverse
library(tidyverse)

# Paste and run the following uncommented code into your console: 
# install.packages("palmerpenguins")

# Save the data as a dataframe
penguins <- as.data.frame(palmerpenguins::penguins)
```

Using a `dplyr` function, pick all the rows/observations in the `penguins` dataset from the year 2007 and save the result as a new object called `penguins_2007`. Compare the number of observations/rows in the original `penguins` dataset with your new `penguins_2007` dataset.

```{r}
#View(penguins)
penguins_2007 <- penguins %>% filter(year == "2007")
#View(penguins_2007)
penguins_obs <- length(penguins$species)
penguins_2007_obs <- length(penguins_2007$species)
print(penguins_obs)
print(penguins_2007_obs)
```

**The penguins_2007 dataset has 110 observations and the penguins dataset has 344 observations.**
------------------------------------------------------------------------

### Question 2: (2 pts)

Using `dplyr` functions on `penguins_2007`, report the number of observations for each species-island combination (note that you'll need to `group_by`). Which species appears on all three islands? 

```{r}
penguins_2007 %>% group_by('species', 'island')
```

**Adelie-Torgersen combination: 20 observations, Adelie-Biscoe combination: 10 observations, Adelie-Dream: 20 observations, Gentoo-Biscoe combination: 34 observations, Chinstrap-Dream combinations: 26 observations. The Adelie species appears on all three islands.**
------------------------------------------------------------------------

### Question 3: (2 pts)

Using `dplyr` functions on `penguins_2007`, create a new variable that contains the ratio of `bill_length_mm` to `bill_depth_mm` (call it `bill_ratio`). *Once you checked that your variable is created correctly*, overwrite `penguins_2007` so it contains this new variable.

```{r}
penguins_2007 <- penguins_2007 %>%
  mutate(bill_ratio = bill_length_mm / bill_depth_mm)
#View(penguins_2007)
```

Are there any cases in the `penguins_2007` dataset for which the `bill_ratio` exceeds 3.5? If so, for which species of penguins is this true?

```{r}
penguin_2007_big_ratio <- penguins_2007 %>% filter(bill_ratio > 3.5)
#View(penguin_2007_big_ratio)
```

**There are 2 observations with bill_ratio > 3.5**
------------------------------------------------------------------------

### Question 4: (2 pts)

Using `dplyr` functions on `penguins_2007`, find the three penguins with the smallest bill ratio for *each species*. Only display the information about `species`, `sex`, and `bill_ratio`. Does the same sex has the smallest bill ratio across species?

```{r}
penguins_2007 %>% group_by(species) %>% arrange(bill_ratio) %>% slice(1:3) %>% select(species, sex, bill_ratio)
```

**For the Adelie and Chinstrap species, the lowest three bill ratios are occupied by the same sex within each species; the lowest three bill ratio values for the Adelie species are all held by males and the lowest three bill ratio values for the Chinstrap species are all held by females. For the Gentoo species, on the other hand, there are both males and females that hold the three lowest bill ratio values.**
------------------------------------------------------------------------

### Question 5: (2 pts)

Using `dplyr` functions on `penguins_2007`, calculate the mean and standard deviation of `bill_ratio` for each species. Drop NAs from `bill_ratio` for these computations (e.g., using the argument `na.rm = T`) so you have values for each species. Which species has the greatest mean `bill_ratio`?

```{r}
penguins_2007 %>% group_by(species) %>% summarize(mean(bill_ratio, na.rm = T), sd(bill_ratio, na.rm = T))
```

**The Gentoo species has the highest mean bill ratio value.**
------------------------------------------------------------------------

### Question 6: (2 pts)

Using `dplyr` functions on `penguins_2007`, remove missing values for `sex`. Pipe a `ggplot` to create a single plot showing the distribution of `body_mass_g` colored by male and female penguins, faceted by species (use the function `facet_wrap()` with the option `nrow = ` to give each species its own row). Which species shows the least sexual dimorphism (i.e., the greatest overlap of male/female size distributions)?

```{r}
# NOT(is.na(sex)) is a vector that holds the negation of whether a value is Nan.
# This means if an observation has a Nan value, then the NOT(is.na(sex)) vector will hold the value FALSE
# With the filter function, all observations with FALSE values in the NOT(is.na(sex)) vector
# will have their sex value removed
penguins_2007 <- penguins_2007 %>% filter(!is.na(sex)) 
ggplot(data=penguins_2007) + geom_histogram(aes(x=body_mass_g, fill=sex)) + facet_wrap(~species, nrow = 3)
```

**The Chinstrap species shows the greatest overlap of body mass between the sexes.**
------------------------------------------------------------------------

### Question 7: (2 pts)

Pipe a `ggplot` to `penguins_2007` to create a scatterplot of `body_mass_g` (y-axis) against `bill_ratio` (x-axis). Does it look like there is a relationship between the bill ratio and the body mass? *Note: you might see a Warning message.* What does this message refer to?*

```{r}
ggplot(data=penguins_2007, aes(x=bill_ratio, y=body_mass_g)) + geom_point()
```

**There appears to be two clusters on the scatter plot. For the cluster with the relatively lower body mass values, there does not seem to be any correlation between bill ratio and body mass, but for the cluster with the higher, relative body mass values, there appears to be a positive correlation between bill ratio and body mass. I did not receive a warning message.**
------------------------------------------------------------------------

### Question 8: (2 pts)

What if we separate each species? Duplicate the plot from the previous question and add a regression trend line with `geom_smooth(method = "lm")`. Color the points and the regression lines by species. Does the relationship between the bill ratio and the body mass changes within each species?

```{r}
ggplot(data=penguins_2007, aes(x=bill_ratio, y=body_mass_g, color=species)) + geom_point() + geom_smooth(method=lm)
```

**For the Adelie species, there is a weak, negative correlation between bill ratio and body mass however for the other two species, there appears to be a very weak, almost nonexistent, positive, correlation between bill ratio and body mass.**
------------------------------------------------------------------------

### Question 9: (2 pts)

Finally, let's make a plot using the original `penguins` dataset (not just the 2007 data). Forewarning: This will be very busy plot!

Map `body_mass_g` to the y-axis, `flipper_length_mm` to the x-axis, `species` to color, and `island` to shape. Using `facet_wrap()`, facet the plots by `year`. Find a way to clean up the x-axis labels (e.g., reduce the amount of tick marks) using `scale_x_continuous()`. Does there appear to be a relationship between body mass and flipper length overall? Is there a relationship within each species? What happens to the distribution of flipper lengths for species over time?

```{r}
ggplot(data=penguins, aes(x=flipper_length_mm, y=body_mass_g, color=species, shape=island)) + geom_point() + facet_wrap(~year) + scale_x_continuous(breaks = seq(175, 225, 15))
```

**There appears to be an overall, strong, positive correlation between flipper length and body mass. There also appears to be a strong, positive correlation between flipper length and body mass within each species. Over time, the overall distribution of flipper length appears to stay relatively the same; however, the distribution of the flipper length for the Gentoo species seems to get a bit larger overtime.**
------------------------------------------------------------------------

### Formatting: (2 pts)

Comment your code, write full sentences, and knit your file!

------------------------------------------------------------------------

```{r, echo=F}
## DO NOT DELETE THIS BLOCK!
Sys.info()
```