---
title: "HW 5"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE,  
                      warning = FALSE, message = FALSE, 
                      fig.align = "center",
                      R.options = list(max.print=100))

# Edit the file starting below
```

### Enter your name and EID here: Callihan Bertley cjb5564

**You will submit this homework assignment as a HTML file on Canvas.**

*For all questions, include the R commands/functions that you used to find your answer (show R chunk). Answers without supporting code will not receive credit. Write full sentences to describe your findings.*

------------------------------------------------------------------------

### Question 1: (1 pt)

The dataset `world_bank_pop` is a built-in dataset in `tidyverse`. It contains information about total population and population growth, overall and more specifically in urban areas, for countries around the world. Take a look at it with `head()`. Is the data tidy? Why or why not?

```{r message=FALSE}
# Call tidyr, dplyr and ggplot2 packages within tidyverse
library(tidyverse)

# Take a look!
head(world_bank_pop)
```

**The dataset is not Tidy because the year variable does not have it's own column; there are multiple variables in a single column.**
------------------------------------------------------------------------

### Question 2: (1 pt)

Using `dplyr` functions on `world_bank_pop`, count how many distinct countries there are in the dataset. Does this makes sense? Why or why not?

```{r}
distinct_countries <- world_bank_pop %>% distinct(country)
print(distinct_countries) 
```

**There are 266 distinct countries in the world_bank_pop data set; however, there are not 266 recognized countries in the world. Even when including regions that are not considered countries, the number of distinct countries seems high.**
------------------------------------------------------------------------

### Question 3: (2 pts)

Use one of the `pivot` functions on `world_bank_pop` to create a new dataset with the years 2000 to 2017 appearing as a *numeric* variable `year`, and the different values for the indicator variable are in a variable called `value`. Save this new dataset in your environment as `myworld1`. 

```{r}
myworld1 <- world_bank_pop %>% pivot_longer(cols = as.character(2000:2017), names_to = 'year', values_to = 'population')
print(myworld1)
```

How many lines are there per country? Why does it make sense?

```{r}
country_ratio <- length(myworld1$population) / length(distinct_countries$country)
print(country_ratio)
```

**There are 72 lines per country. This makes sense because there are 18 different values for year and 4 differnt values for the indicator column.**
------------------------------------------------------------------------

### Question 4: (3 pts)

Use another `pivot` function on `myworld1` to create a new dataset, `myworld2`, with the different categories for the indicator variable appearing as their own variables. Use `dplyr` functions to rename `SP.POP.GROW` and `SP.URB.GROW`, as `pop_growth` and `pop_urb_growth` respectively. 

```{r}
#View(myworld1) 

myworld2 <- myworld1 %>% pivot_wider( names_from = indicator, values_from = population)

myworld2 <- myworld2 %>% mutate(pop_growth = `SP.POP.GROW`, pop_urb_growth = `SP.URB.GROW`) %>% select(-`SP.POP.GROW`, -`SP.URB.GROW`)

#View(myworld2)
```

Using `dplyr` functions, find the ratio of urban growth compared to the population growth in the world for each year. *Hint: the country code `WLD` represents the entire world.* Create a `ggplot` to display how the percentage of urban population growth has changed over the years. Why does your graph not contradict the fact that the urban population worldwide is increasing over the years?

```{r}
myworld3 <- myworld2 %>% mutate(urb_growth_pop_growth_ratio = pop_urb_growth / pop_growth)
world_df <- myworld3 %>% filter(country == 'WLD')
#View(world_df)
ggplot(world_df, aes(year, urb_growth_pop_growth_ratio)) + geom_point() + xlab("Country") + ylab("Urban Growth to Population Growth Ratio") + ggtitle("Urban Growth to Population Growth Ratio Over Time")
```

**The plot does not contradict that the urban population is increasing over the years because a ratio of greater than 1 implies that the urban growth rate is increasing at a higher rate than the rate at which the world's population is increasing.**
------------------------------------------------------------------------

### Question 5: (1 pt)

In `myworld2`, which country code had the highest population growth in 2017?

```{r}
myworld2 %>% arrange(desc(pop_growth)) %>% filter(year==2017)
```

**Qatar had the highest population growth 2017.**
------------------------------------------------------------------------

### Question 6: (1 pt)

When answering the previous, we only reported the three-letter code and (probably) have no idea what the actual country is. We will now use the package `countrycode` with a built-in dataset called `codelist` that has information about the coding system used by the World bank:

```{r}
# Paste and run the following into your console (NOT HERE): install.packages("countrycode")

# Call the countrycode package
library(countrycode)

# Create a list of codes with matching country names
mycodes <- codelist

print(mycodes)
```

Using `dplyr` functions, modify `mycodes` above to only keep the variables `continent`, `wb` (World Bank code), and `country.name.en` (country name in English). Then remove countries with missing `wb` code.

How many countries are there in `mycodes`? 

```{r}
 mycodes <- codelist %>% select(continent, country.name.en, wb)

mycodes <- mycodes %>% filter(!(is.na(wb)))

mycodes %>% distinct(country.name.en)
```

**There are unique 218 countries in mycodes**
------------------------------------------------------------------------

### Question 7: (1 pt)

Use a `left_join()` function to add the information of the country codes **to** `myworld2` dataset. Match the two datasets based on the World Bank code. *Note: the World Bank code does not have the same name in each dataset.* Using `dplyr` functions, only keep the data available for Europe and for the year 2017. Save this new dataset as `myeurope`.

```{r}
myworld4 <- left_join(myworld2, mycodes, by = c('country' = 'wb'))

myeurope <- myworld4 %>% filter(year == 2017, continent == "Europe")
print(myeurope)
```

How many rows are there in `this new dataset`myeurope`? What does each row represent?

```{r}
dim(myeurope)
myeurope
```

**There are 46 rows in myeurope where each row represents a european country in 2017.**
------------------------------------------------------------------------

### Question 8: (2 pts)

Using `dplyr` functions on `myeurope`, only keep information for the population growth in 2017 then compare the population growth per country with `ggplot` using `geom_bar()`. Make sure to order countries in order of population growth. Which country in Europe had the lowest population growth in 2017?

```{r}
myeuropop <- myeurope %>% select(country, country.name.en, pop_growth) %>% arrange(desc(pop_growth))

myeuropop$country <- factor(myeuropop$country, levels = myeuropop$country[order(myeuropop$pop_growth)])

ggplot(myeuropop, aes(country, pop_growth)) + geom_bar(stat = 'identity') + coord_flip() + theme_minimal()
```

**Moldova had the lowest population growth in 2017**
------------------------------------------------------------------------

### Question 9: (1 pt)

When dealing with location data, we can actually visualize information on a map if we have geographic information such as latitude and longitude. Next, we will use a built-in function called `map_data()` to get geographic coordinates about countries in the world (see below). Take a look at the dataset `mapWorld`. What variables could we use to join `mapWorld` and `myeurope`? *Note: the variables do not have the same name in each dataset but they contain the same information.*

```{r}
# Geographic coordinates about countries in the world
mapWorld <- map_data("world")
mapWorld
myeurope
mapWorld %>% filter(region == 'France')
```

**We could join the two datasets on the region column of mapWorld and the country.name.en column of myeurope.**
------------------------------------------------------------------------

### Question 10: (2 pts)

Use a joining function to check if any information from `myeurope` is not contained in `mapWorld`, matching the two datasets based on the country name.

```{r}
anti_join(myeurope, mapWorld, by = c("country.name.en" = "region"))
```

Some countries such as United Kingdom did not have a match. Why do you think this happened? *Hint: find the distinct country names in `mapWorld`, arrange them in alphabetical order, and scroll through the names. Can you find any of these countries with no match in a slightly different form?*

```{r}
mapWorld %>% distinct(region) %>% arrange()
```

**Some of the regions in myMap are not found in myeurope because they are written differently in the two datasets. I was able to find the UK, the Czech Republican, and Bosnia & Herzegovina in the myMap dataset with their names written differently however I was not able to find Gibraltar. I think that because it is not an independent country it was not listed in the myMap dataset.**
------------------------------------------------------------------------

### Question 11: (1 pt)

Consider the `myeurope` dataset. Recode some of the country names so that the countries with no match from the previous question (with the exception of Gibraltar which is not technically a country anyway) will have a match. *Hint: use `recode()` inside `mutate()` as described in this article https://www.statology.org/recode-dplyr/.* Then add a pipe and use a `left_join()` function to add the geographic information in `mapWorld` to the countries in `myeurope`. Save this new dataset as `mymap`.

```{r}
# your code goes below (make sure to edit comment)
myeurope <- myeurope %>% mutate(country.name.en = recode(country.name.en,
                                        "Bosnia & Herzegovina" = "Bosnia and Herzegovina",
                                        "United Kingdom" = "UK",
                                        "Czechia" = "Czech Republic"))



mymap <- left_join(myeurope, mapWorld, by = c("country.name.en" = "region"))
```

------------------------------------------------------------------------

### Question 12: (2 pts)

Let's visualize how population growth varies across European countries in 2017 with a map. With the package `ggmap`, use the R code provided below. Add a comment after each `#` to explain what each component of this code does. *Note: it would be a good idea to run the code piece by piece to see what each layer adds to the plot.*

```{r}
# Paste and run the following into your console (NOT HERE): install.packages("ggmap")

# Call the ggmap package
library(ggmap)

# Build a map!
mymap %>%
  # Access the mymap dataset built into the ggmap library
  ggplot(aes(x = long, y = lat, group = group, fill = pop_growth)) + 
  # Create a plot of the longitute and latitude, split by group and 
  # colored depending on population growth
  geom_polygon(colour = "black") +
  # Color the country outlines
  scale_fill_gradient(low = "white", high = "blue") +
  # Creates a coloring gradient where the countries with lower population
  # growth are more white and the countries with higher population growth are more blue
  labs(fill = "Growth" ,title = "Population Growth in 2000", 
       x ="Longitude", y ="Latitude") +
  # Adds labels to the plot
  xlim(-25,50) + ylim(35,70)
  # Creates bounds on the longitude and latitude on the plot
```

Which country had the highest population growth in Europe in 2017? *Hint: it's very tiny and very close to where I'm from! You can refer to this map for European geography: https://www.wpmap.org/europe-map-hd-with-countries/*

**It appears that Malta had the highest growth rate in 2017 compared to all other European countries.**
------------------------------------------------------------------------

### Formatting: (2 pts)

Comment your code, write full sentences, and knit your file!

------------------------------------------------------------------------

```{r, echo=F}
## DO NOT DELETE THIS BLOCK!
Sys.info()
```