---
title: "HW 2"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE,  
                      warning = FALSE, message = FALSE, 
                      fig.align = "center",
                      R.options = list(max.print=100))

# Edit the file starting below
```

### Enter your name and EID here: Callihan Bertley cjb5564

**You will submit this homework assignment as as HTML file on Canvas. PDF or .Rmd submissions receive zero.**

*For all questions, include the R commands/functions that you used to find your answer (show R chunk). Answers without supporting code will not receive credit. Write full sentences to describe your findings.*

**The goal of this assignment is to encode your name (or any other message) using a *cipher* function: We want to replace each letter of a given character vector with the letter of the alphabet that is *k* positions after it in the alphabet. For example, if the letter was `a` and `k = 3`, we would replace it with `d`. We will also want it to loop around, so if the letter was `y` and `k = 3`, we'd replace it with `b`. For example, with `k = 3`, the word `dog` would become `grj`. Let's take it step by step.**

------------------------------------------------------------------------

### Question 1: (2 pts)

Type the word `letters` into the R chunk below. What does this predefined object in R contain? What is this object's data type/class? How many elements does it contain? *Include base R commands used to answer all three questions.*

```{r}
letters
typeof(letters)
length(letters)
```
**'letters' is an object that contains all lowercase, English letters and is of datatype character. There are 26 elements in the 'letters' object.  **
------------------------------------------------------------------------

### Question 2: (2 pts)

First, here is the code to split a word into a vector containing each letter.

```{r}
test <- unlist(strsplit("test", split = ""))
# Note: the function strsplit() returns a list, use unlist() to return a vector
```

Remember that `A %in% B` returns a vector of the positions of matches of an object A in an object B (Worksheet 2):

```{r}
letters %in% test
```

How many elements are false in the resulting logical vector? *Include base R commands used to answer this question (recall: TRUE is equivalent to the value 1 and FALSE is 0.*

```{r}
logical_vec <-letters %in% test
print(length(logical_vec[logical_vec == FALSE]))
```
**The logical vector created from the command "letters %in% test" holds 23 False values. **
------------------------------------------------------------------------

### Question 3: (2 pts)

Another function that will be useful is `which()`: it takes a logical vector as an input and returns the indices/positions that are TRUE. For example, run the following code:

```{r}
# Note: T is shorthand for TRUE, F for FALSE
which(c(F,T,F,T,F,T))
```

The output tells you that elements in position 2, 4, and 6 are true. Now, use the `which` function, along with `%in%` and `letters`, to find which positions in the alphabet the letters in the name `arya` occupy (saved as an object called `name_v` below). Would that combination of functions alone work to encode a name? Why/Why not?

```{r}
# Define name as a vector
name_v <- unlist(strsplit("arya", split = ""))

# Which letters are in the name
which(letters %in% name_v)
```
**We could not encode a name with this technique because it only returns each letter once, so many names with multiple of the same letter could not be properly deciphered or decoded with full certainty.**
------------------------------------------------------------------------

### Question 4: (2 pts)

How can we avoid this? For example, we can test each letter one at a time in their correct order! One approach would be to use a *for loop*. Write a for loop that goes through each element of the character vector `name_v` (i.e., each letter in `c("a","r","y","a")`) one at a time, finds its position in the alphabet, and saves each position in a vector called `positions`. Confirm that the positions are correct by using `positions` as an index to find the corresponding letters in the object `letters`.

*For example, the name `ali` would give you the positions 1,12, and 9. You can grab the letters in those positions by doing* `letters[c(1, 12, 9)]`.

```{r}
positions <- c()
for(i in 1: length(letters)){
  positions <- append(positions, (which(letters %in% name_v[i])))
}
positions
letters[positions]
```

------------------------------------------------------------------------

### Question 5: (2 pts)

Let's encode the name `arya`! Shift all the `positions` by 1 and index `letters` to obtain the encoded name. Is the encoded name a real name?

```{r}
for(i in 1:length(positions)){
  positions[i] = positions[i] + 1
}
letters[positions]

```
**After shifting the position vector by 1, the name "bszb" is printed. The name "bszb" is not a real name.**
------------------------------------------------------------------------

### Question 6: (2 pts)

Now, what if you would like to get the positions in your name? Or any other name? We would have to repeat questions 2-5... Instead let's write a function to 1) split a name as a vector (i.e., a character vector whose elements contain single letters), 2) initialize the positions, and 3) report each position in a vector `positions` with a for loop for each new name we would like to encode. The function should take a `name` (for example, "layla") as the input and return the alphabetical positions each of those letters occupy. Call the function `get_position`. Once you have defined it, test it out with "layla". Did you get all positions?

```{r}
get_position <- function(name){
  name_vec <- unlist(strsplit(name, split = ""))
  positions <- c()
  for(i in 1:length(letters)){
    positions <-append(positions, which(letters %in% name_vec[i]))
  }
  print(positions)
  print(letters[positions])
}

get_position('layla')
```
**When using the get_position function for the name 'layla' the positions returned were correct; the function returned all of the letter positions.**
------------------------------------------------------------------------

### Question 7: (2 pts)

What happens when we shift the positions past `z`, the 26th and final letter of the alphabet? Shifting the positions in `arya` up by `k = 2` should give `ncanc`, but since there is no 27th element of letters, it will return `NA` instead of `a`. Try it in the code chunk below.

```{r}
#letters[get_position(name_v) + 2] (commented out this code so the file could knit)
# returns NA instead of a 
```

How do we make it loop around so that `z` shifted up 1 becomes `a`? In other words, how can we make 27 become 1, 28 become 2, 29 becomes 3, etc.? We will use a mathematical operator called modulo `%%` (which tells you the remainder when you divide one number by another). Try running the code below, `27 %% 26` (pronounced "27 modulo 26"). It returns 1, the remainder when the number on the left (27) is divided by the number on the right (26).

```{r}
27 %% 26
```

We just need our shifted positions *modulo* 26. You can do this with `(positions + k) %% 26`. One last minor issue: `26 %% 26` is 0 (or any multiple of 26 %% 26 is 0) but we want it to return 26 (i.e., the letter `z`). We can fix this issue by using `ifelse` for example. Test if `positions + k %% 26` is 0: if it is, use 26, if it is not use `positions + k %% 26`. Use your function `get_position()` and the fix with modulo `%%` in `ifelse()` to encode the word `arya` by shifting every letter `k = 2` positions forward correctly. Is the encoded name a real name?

```{r}
get_position <- function(name){
  name_vec <- unlist(strsplit(name, split = ""))
  positions <- c()
  for(i in 1:length(letters)){
    positions <-append(positions, which(letters %in% name_vec[i]))
  }
  k <- 2
  positions <- ifelse((positions + k) %% 26 == 0, positions + k, (positions + k) %% 26)
  print(positions)
  print(letters[positions])
}

get_position('arya')
```
**The output vector, once shifted by 2 units, corresponds to the name 'ctac' which is not a real name.**
------------------------------------------------------------------------

### Question 8: (2 pts)

Putting it all together: Write a function that incorporates all the work you have done to achieve the encoding task. Name the function `cipher`. This function should take two arguments: a `name` (a string) and how many positions to shift (`k`). Fill in the code below with what you have been using above. Check your code with `arya` shifted by 2 positions and test your code with your own name with the shift of your choice! Is the encoded name a real name?

```{r}
cipher <- function(name, k) {
  
  name_v <- unlist(strsplit(name, split = ""))
  
  positions <- c()
  for (i in 1:length(letters)) {
    positions <-append(positions, which(letters %in% name_v[i]))
  }
  positions <- ifelse((positions + k) %% 26 == 0, positions + k, (positions + k) %% 26)
  print(positions)
  print(letters[positions])
}

# check
cipher("arya", 2)

# test with your name!
cipher("callihan", 5)

```
**My name ("callihan") shifted by 5 letters, once ciphered, is "hfqqnmfs" which is not a real name.**
------------------------------------------------------------------------

### Question 9: (2 pts)

A less guided question... You were given the code `oldp`. Can you decipher the code and find the name hidden behind it?

```{r}
decipher <- function(name, k){
  
  name_v <- unlist(strsplit(name, split = ""))
  
  positions <- c()
  for (i in 1:length(letters)) {
    positions <-append(positions, which(letters %in% name_v[i]))
  }
  positions <- ifelse((positions - k) %% 26 == 0, positions -k + 26, (positions - k) %% 26)
  print(positions)
  print(letters[positions])
}

# All possible decoded names
for(k in 1:26){
  decipher('oldp',k)
}
# Potential Names
decipher('oldp', 3) # liam
```
**After creating a function to decipher the name and running a for loop with k values (here k is the number of units to shift in the reverse order to the cipher code) in the range [1, 26] to view all possible decoded names, I have reached the name "liam" when shifting 3 units back (k=3). The name is "liam".**
------------------------------------------------------------------------

### Formatting: (2 pts)

Comment your code, write full sentences, and knit your file!

------------------------------------------------------------------------

```{r, echo=F}
## DO NOT DELETE THIS BLOCK!
Sys.info()
```
